\documentclass[12pt]{article} 

\begin{document}

\title{Lab 5 Assignment [Ushna Ijaz- 2019-CE-39]}
\maketitle

\section*{Question 1}
Suppose we are given an array A of length n with the promise that
there exists a majority element (i.e. an element that appears > n / 2 times). Additionally,
we are only allowed to check whether two elements are equal (no > or < comparisons).
Design an $O(nlog(n))$ algorithm to find the majority element, using divide and conquer.
Informally, explain the correctness and runtime of your algorithm. [We are expecting
actual code, an English description of the main idea of the algorithm, as well as an informal
explanation of correctness and runtime] \\



\section*{Solution}

We split the array A into 2 subarrays arr1 and arr2 of half the size. We choose the majority element of arr1
and arr2. The recurrence:\\

$T(n) = 2T(n/2)+O(n)$ \\

We then use the divide-and-conquer method. Basically we have to pair up the elements so we get $n/2$. If the two elements in a pair are difference, it is discarded.
If one of the parts has a majority element we count the  number of times it is repeated. And if both parts have the same majority element, it is automatically the majority element for array A. My code counts the occurrences for the majority elements in each half of the list. \\


\section*{Question 2}

For the sorted arrays, we count while we compare the elements in the given arrays. But first we need to merge both the arrays. For that we took help from the Merge Sort algorithm. Its basically finds the medians of the two sorted arrays and second step is that it compares them.When count becomes n we got our answer. For the Unsorted arrays, the procedure is almost the same but the only difference made is that I sorted the arrays.  




\section*{Question 3 Table}



\begin{tabular}{ |p{3cm}||p{3cm}|p{3cm}|p{3cm}|  }
 \hline
 \multicolumn{4}{|c|}{Time Comparison} \\
 \hline
 Array Size & Insertion Sort & Merge Sort & Quick Sort\\
 \hline
 10   & 0.0    &0.0&   0.0\\
 100 &   0.0  & 0.0   &0.007989883\\
 1000 &0.124124526 & 0.0100557804&  0.010042428\\
 10000    &15.5295608 & 0.08614063&   0.56423211\\
 100000&   1788.871452  & 2.18674635&29.0143427\\

 \hline
\end{tabular}







\end{document}
